---
title: "ESG Project"
author: "Karene Matoka"
date: '2022-05-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(broom)
library(nlme)
library(broom.mixed) 
library(GGally)
library(car)
```

# TIDY DATA

```{r, message = FALSE}
# CHANGE THIS TO BE THE CORRECT UPLOAD PATH FOR YOUR COMPUTER
esg <- read_csv("./ESGData.csv.csv")
esg %>%
  rename(Variable = "Indicator Name") %>%
  rename("Variable code" = "Indicator Code") -> esg

countries <- c("Afghanistan", "Albania", "Algeria", "Andorra", "Angola", "Antigua and Barbuda", "Argentina", "Armenia", "Australia", "Austria", "Azerbaijan", "Bahamas, The", "Bahrain", "Bangladesh", "Barbados", "Belarus", "Belgium", "Belize", "Benin", "Bhutan", "Bolivia", "Bosnia and Herzegovina", "Botswana", "Brazil", "Brunei Darussalam", "Bulgaria", "Burkina Faso", "Burundi", "Cabo Verde", "Cambodia", "Cameroon", "Canada", "Central African Republic", "Chad", "Chile", "China", "Colombia", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Costa Rica", "Cote d'Ivoire", "Croatia", "Cuba", "Cyprus", "Czech Republic", "Denmark", "Djibouti", "Dominica", "Dominican Republic", "Ecuador", "Egypt, Arab Rep.", "El Salvador", "Equatorial Guinea", "Eritrea", "Estonia", "Eswatini", "Ethiopia", "Fiji", "Finland", "France", "Gabon", "Gambia, The", "Georgia", "Germany", "Ghana", "Greece", "Grenada", "Guatemala", "Guinea", "Guinea-Bissau", "Guyana", "Haiti", "Honduras", "Hungary", "Iceland", "India", "Indonesia", "Iran, Islamic Rep.", "Iraq", "Ireland", "Israel", "Italy", "Jamaica", "Japan", "Jordan", "Kazakhstan", "Kenya", "Kiribati", "Korea, Dem. People's Rep.", "Korea, Rep.", "Kuwait", "Kyrgyz Republic", "Lao PDR", "Latvia", "Lebanon", "Lesotho", "Liberia", "Libya", "Liechtenstein", "Lithuania", "Luxembourg", "Madagascar", "Malawi", "Malaysia", "Maldives", "Mali", "Malta","Marshall Islands", "Mauritania", "Mauritius", "Mexico", "Micronesia, Fed. Sts.", "Moldova", "Monaco", "Mongolia", "Montenegro", "Morocco", "Mozambique", "Myanmar", "Namibia", "Nauru", "Nepal", "Netherlands", "New Zealand", "Nicaragua", "Niger", "Nigeria", "North Macedonia", "Norway", "Oman", "Pakistan", "Palau", "Panama", "Papua New Guinea", "Paraguay", "Peru", "Philippines", "Poland", "Portugal", "Qatar", "Romania", "Russian Federation", "Rwanda", "Samoa", "San Marino", "Sao Tome and Principe", "Saudi Arabia", "Senegal", "Serbia", "Seychelles", "Sierra Leone", "Singapore", "Slovak Republic", "Slovenia", "Solomon Islands", "Somalia", "South Africa", "South Sudan", "Spain", "Sri Lanka", "St. Kitts and Nevis", "St. Lucia", "St. Vincent and the Grenadines", "Sudan", "Suriname", "Sweden", "Switzerland", "Syrian Arab Republic", "Tajikistan", "Tanzania", "Thailand", "Timor-Leste", "Togo", "Tonga","Trinidad and Tobago", "Tunisia", "Turkey", "Turkmenistan", "Tuvalu", "Uganda", "Ukraine", "United Arab Emirates", "United Kingdom", "United States", "Uruguay", "Uzbekistan", "Vanuatu", "Venezuela, RB", "Vietnam", "Yemen, Rep.", "Zambia", "Zimbabwe")

# countries <- c("United States", "Canada", "Somalia", "Germany", "Spain", "France", "United Arab Emirates", "Haiti", "Brazil", "India", "Australia", "South Africa", "Japan", "Mexico", "Iceland", "Cuba")
esg %>%
  filter(`Country Name` %in% countries) -> esg_countries
esg_countries
```

```{r, message=FALSE}
esg_countries %>%
  filter(Variable == "Life expectancy at birth, total (years)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2019`) %>%
    pivot_longer(
    cols = `1960`:`2019`,
    names_to = "Year",
    values_to = "Life_Exp"
  ) %>%
  select(`Country Name`, `Year`, `Life_Exp`) -> lifeexp

# DROUGHT 
esg_countries %>%
  filter(Variable == "Droughts, floods, extreme temperatures (% of population, average 1990-2009)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `2009`) %>%
      pivot_longer(
     cols = `2009`,
     names_to = "Year",
     values_to = "Drought"
   ) %>%
   select(`Country Name`, `Year`, `Drought`) -> drought

# FOSSIL FUELS
 esg_countries %>%
   filter(Variable == "Fossil fuel energy consumption (% of total)") %>%
   select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2015`) %>%
      pivot_longer(
     cols = `1960`:`2015`,
     names_to = "Year",
     values_to = "Fossil_Fuels"
   ) %>%
   select(`Country Name`, `Year`, `Fossil_Fuels`) -> fossilfuels

# CO2 EMISSIONS
esg_countries %>%
  filter(Variable == "CO2 emissions (metric tons per capita)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2018`) %>%
      pivot_longer(
     cols = `1960`:`2018`,
     names_to = "Year",
     values_to = "CO2"
   ) %>%
   select(`Country Name`, `Year`, `CO2`) -> co2

# OVERWEIGHT
esg_countries %>%
  filter(Variable == "Prevalence of overweight (% of adults)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1975`:`2016`) %>%
      pivot_longer(
     cols = `1975`:`2016`,
     names_to = "Year",
     values_to = "Overweight"
   ) %>%
   select(`Country Name`, `Year`, `Overweight`)-> overweight 

# UNDER NOURISHMENT
esg_countries %>%
  filter(Variable == "Prevalence of undernourishment (% of population)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `2001`:`2019`) %>%
      pivot_longer(
     cols = `2001`:`2019`,
     names_to = "Year",
     values_to = "Undernourishment"
   ) %>%
   select(`Country Name`, `Year`, `Undernourishment`) -> underweight 

# HOSPITAL BEDS
esg_countries %>%
  filter(Variable == "Hospital beds (per 1,000 people)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2019`) %>%
      pivot_longer(
     cols = `1960`:`2019`,
     names_to = "Year",
     values_to = "Hospital_Beds"
   ) %>%
   select(`Country Name`, `Year`, `Hospital_Beds`) -> hospitalbed 

# MORTALITY RATE UNDER 5
esg_countries %>%
   filter(Variable == "Mortality rate, under-5 (per 1,000 live births)") %>%
   select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2019`) %>%
      pivot_longer(
     cols = `1960`:`2019`,
     names_to = "Year",
     values_to = "Mortality_Rate"
   ) %>%
   select(`Country Name`, `Year`, `Mortality_Rate`)-> mortality

# FERTILITY RATE
esg_countries %>%
  filter(Variable == "Fertility rate, total (births per woman)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1960`:`2019`) %>%
      pivot_longer(
     cols = `1960`:`2019`,
     names_to = "Year",
     values_to = "Fertility_Rate"
   ) %>%
   select(`Country Name`, `Year`, `Fertility_Rate`)-> fertility

# GOV EFFFECTIVENESS
esg_countries %>%
  filter(Variable == "Government Effectiveness: Estimate") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1996`:`2019`) %>%
      pivot_longer(
     cols = `1996`:`2019`,
     names_to = "Year",
     values_to = "Government_Effectiveness"
   ) %>%
   select(`Country Name`, `Year`, `Government_Effectiveness`)-> goveffective

# WOMEN IN PARLIAMENT
esg_countries %>%
  filter(Variable == "Proportion of seats held by women in national parliaments (%)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1997`:`2019`) %>%
      pivot_longer(
     cols = `1997`:`2019`,
     names_to = "Year",
     values_to = "Women_Parliament"
   ) %>%
   select(`Country Name`, `Year`, `Women_Parliament`) -> womenparliament

# POLITICAL STABILITY
esg_countries %>%
  filter(Variable == "Political Stability and Absence of Violence/Terrorism: Estimate") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1996`:`2019`) %>%
      pivot_longer(
     cols = `1996`:`2019`,
     names_to = "Year",
     values_to = "Political_Stability"
   ) %>%
   select(`Country Name`, `Year`, `Political_Stability`)-> politicstability

# RULE OF LAW
esg_countries %>%
  filter(Variable == "Rule of Law: Estimate") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1996`:`2019`) %>%
      pivot_longer(
     cols = `1996`:`2019`,
     names_to = "Year",
     values_to = "Rule_Law"
   ) %>%
   select(`Country Name`, `Year`, `Rule_Law`) -> ruleoflaw

# LEGAL RIGHTS INDEX
esg_countries %>%
  filter(Variable == "Strength of legal rights index (0=weak to 12=strong)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `2013`:`2019`) %>%
      pivot_longer(
     cols = `2013`:`2019`,
     names_to = "Year",
     values_to = "Legal_Rights"
   ) %>%
   select(`Country Name`, `Year`, `Legal_Rights`) -> legalrights

# GDP
esg_countries %>%
  filter(Variable == "GDP growth (annual %)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1961`:`2019`) %>%
      pivot_longer(
     cols = `1961`:`2019`,
     names_to = "Year",
     values_to = "GDP"
   ) %>%
   select(`Country Name`, `Year`, `GDP`)-> gdp

# SCHOOL ENROLLMENT
esg_countries %>%
  filter(Variable == "School enrollment, primary and secondary (gross), gender parity index (GPI)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1971`:`2019`) %>%
      pivot_longer(
     cols = `1971`:`2019`,
     names_to = "Year",
     values_to = "School_Enrollment"
   ) %>%
   select(`Country Name`, `Year`, `School_Enrollment`)-> school

# LABOR FORCE
esg_countries %>%
  filter(Variable == "Ratio of female to male labor force participation rate (%) (modeled ILO estimate)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1990`:`2019`) %>%
      pivot_longer(
     cols = `1990`:`2019`,
     names_to = "Year",
     values_to = "Women_LaborForce"
   ) %>%
   select(`Country Name`, `Year`, `Women_LaborForce`)-> labor

# VOICE AND ACCOUNTABILITY
esg_countries %>%
  filter(Variable == "Voice and Accountability: Estimate") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1996`:`2019`) %>%
      pivot_longer(
     cols = `1996`:`2019`,
     names_to = "Year",
     values_to = "Voice_Accountability"
   ) %>%
   select(`Country Name`, `Year`, `Voice_Accountability`) -> voice

# INTERNET USE
esg_countries %>%
  filter(Variable == "Individuals using the Internet (% of population)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1990`:`2019`) %>%
      pivot_longer(
     cols = `1990`:`2019`,
     names_to = "Year",
     values_to = "Internet_Use"
   ) %>%
   select(`Country Name`, `Year`, `Internet_Use`) -> internet

# POPULATION DENSITY
esg_countries %>%
  filter(Variable == "Population density (people per sq. km of land area)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1961`:`2019`) %>%
      pivot_longer(
     cols = `1961`:`2019`,
     names_to = "Year",
     values_to = "Population_Density"
   ) %>%
   select(`Country Name`, `Year`, `Population_Density`)-> density

# LITERACY RATE
esg_countries %>%
  filter(Variable == "Literacy rate, adult total (% of people ages 15 and above)") %>%
  select(`Country Name`, `Country Code`, Variable, `Variable code`, `1972`:`2019`) %>%
      pivot_longer(
     cols = `1972`:`2010`,
     names_to = "Year",
     values_to = "Literacy_Rate"
   ) %>%
   select(`Country Name`, `Year`, `Literacy_Rate`) -> literacy
```

```{r, message = FALSE}
left_join(lifeexp, drought, by = c("Year", "Country Name")) -> df1
left_join(df1, fossilfuels, by = c("Year", "Country Name")) -> df2 
left_join(df2, co2, by = c("Year", "Country Name")) -> df3 
left_join(df3, overweight, by = c("Year", "Country Name")) -> df4 
left_join(df4, underweight, by = c("Year", "Country Name")) -> df5 
left_join(df5, hospitalbed, by = c("Year", "Country Name")) -> df6 
left_join(df6, mortality, by = c("Year", "Country Name")) -> df7 
left_join(df7, fertility, by = c("Year", "Country Name")) -> df8 
left_join(df8, goveffective, by = c("Year", "Country Name")) -> df9 
left_join(df9, womenparliament, by = c("Year", "Country Name")) -> df10 
left_join(df10, politicstability, by = c("Year", "Country Name")) -> df11
left_join(df11, ruleoflaw, by = c("Year", "Country Name")) -> df12 
left_join(df12, legalrights, by = c("Year", "Country Name")) -> df13 
left_join(df13, gdp, by = c("Year", "Country Name")) -> df14 
left_join(df14, school, by = c("Year", "Country Name")) -> df15 
left_join(df15, labor, by = c("Year", "Country Name")) -> df16 
left_join(df16, voice, by = c("Year", "Country Name")) -> df17 
left_join(df17, internet, by = c("Year", "Country Name")) -> df18
left_join(df18, density, by = c("Year", "Country Name")) -> df19
left_join(df19, literacy, by = c("Year", "Country Name")) -> esg 
esg %>%
  rename(Country = "Country Name") %>%
  mutate(Year = as.numeric(Year)) %>%
  filter(!is.na(Life_Exp)) %>%
  mutate(log_lifeexp = log(Life_Exp)) -> esg
head(esg)

```

```{r ESGSUB}
map_dbl(esg, ~mean(is.na(.)))

nrow(esg)

esg %>%
  discard(map_dbl(esg, ~mean(is.na(.))) > 0.7) %>%
  filter(!is.na(Country),
         !is.na(Year),
         !is.na(Life_Exp),
         !is.na(Fossil_Fuels),
         !is.na(CO2),
         !is.na(Overweight),
         !is.na(Fertility_Rate),
         !is.na(GDP),
         !is.na(Population_Density),
         !is.na(School_Enrollment),
         !is.na(Women_LaborForce),
         !is.na(Internet_Use),
         !is.na(Hospital_Beds),
         !is.na(Government_Effectiveness),
         !is.na(Women_Parliament),
         !is.na(Political_Stability),
         !is.na(Rule_Law),
         !is.na(School_Enrollment),
         !is.na(Voice_Accountability)) %>%
  select(-"Mortality_Rate") ->
  esg_sub
head(esg_sub)

lmall <- lme(fixed = log_lifeexp ~ Fossil_Fuels + CO2 + Overweight + Fertility_Rate + GDP + Population_Density + School_Enrollment + Women_LaborForce + Internet_Use + Hospital_Beds + Government_Effectiveness + Women_Parliament + Political_Stability + Rule_Law + Voice_Accountability, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_sub)
tidy(lmall)
plot(lmall)
```

# EXPLORATORY DATA ANALYSIS

```{r EDA}
ggpairs(esg_sub, columns = 3:8)
ggpairs(esg_sub, columns = c(3, 9:13))
ggpairs(esg_sub, columns = c(3, 14:18))
```

```{r}
esg_sub %>%
  filter(!Fossil_Fuels == 0) %>%
  filter(!Women_Parliament == 0) %>%
  mutate(log_ff = log(Fossil_Fuels)) %>%
  mutate(log_co2 = log(CO2)) %>%
  mutate(log_overweight = log(Overweight)) %>%
  mutate(log_fertility = log(Fertility_Rate)) %>%
  mutate(log_pop = log(Population_Density)) %>%
  mutate(log_school = log(School_Enrollment)) %>%
  mutate(log_wlabor = log(Women_LaborForce)) %>%
  mutate(log_internet = log(Internet_Use)) %>%
  mutate(log_beds = log(Hospital_Beds)) %>%
  mutate(log_wparliament = log(Women_Parliament))  -> esg_logs
head(esg_logs)
```



# SIMPLE LINEAR REGRESSION & RESIDUAL PLOTS

```{r, message= FALSE}
# FOSSIL FUELS
lmff <- lme(fixed = log_lifeexp ~ Fossil_Fuels, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmff)

aff <- augment(lmff)
qplot(x = .fitted, y = .resid, data = aff) + geom_hline(yintercept = 0)
```
```{r }
# CO2
lmco2 <- lme(fixed = log_lifeexp ~ log_co2, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmco2)

aco2 <- augment(lmco2)
qplot(x = .fitted, y = .resid, data = aco2) + geom_hline(yintercept = 0)
```
```{r }
# OVERWEIGHT
lmoverweight <- lme(fixed = log_lifeexp ~ Overweight, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmoverweight)

aoverweight <- augment(lmoverweight)
qplot(x = .fitted, y = .resid, data = aoverweight) + geom_hline(yintercept = 0)
```
```{r}
# HOSPITAL BEDS
lmhosp <- lme(fixed = log_lifeexp ~ log_beds, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmhosp)

ahosp <- augment(lmhosp)
qplot(x = .fitted, y = .resid, data = ahosp) + geom_hline(yintercept = 0)
```
```{r }
# FERTILITY RATE
lmfertility <- lme(fixed = log_lifeexp ~ log_fertility, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmfertility)

afertility <- augment(lmfertility)
qplot(x = .fitted, y = .resid, data = afertility) + geom_hline(yintercept = 0)
```
```{r}

# GOV EFFECTIVENESS 
lmgov <- lme(fixed = log_lifeexp ~ Government_Effectiveness, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmgov)

agov <- augment(lmgov)
qplot(x = .fitted, y = .resid, data = agov) + geom_hline(yintercept = 0)
```
```{r}

# WOMEN IN PARLIAMENT
lmwomen <- lme(fixed = log_lifeexp ~ log_wparliament, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmwomen)

awomen <- augment(lmwomen)
qplot(x = .fitted, y = .resid, data = awomen) + geom_hline(yintercept = 0)
```
```{r }
# POLITICAL STABILITY
lmpolitical <- lme(fixed = Life_Exp ~ Political_Stability, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmpolitical)

apolitical <- augment(lmpolitical)
qplot(x = .fitted, y = .resid, data = apolitical) + geom_hline(yintercept = 0)
```
```{r }
# RULE OF LAW 
lmrule <- lme(fixed = Life_Exp ~ Rule_Law, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmrule)

arule <- augment(lmrule)
qplot(x = .fitted, y = .resid, data = arule) + geom_hline(yintercept = 0)
```
```{r }
# GDP 
lmgdp <- lme(fixed = log_lifeexp ~ GDP, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmgdp)

agdp <- augment(lmgdp)
qplot(x = .fitted, y = .resid, data = agdp) + geom_hline(yintercept = 0)
```
```{r }
# SCHOOL ENROLLMENT
lmschool <- lme(fixed = log_lifeexp ~ School_Enrollment, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmschool)

aschool <- augment(lmschool)
qplot(x = .fitted, y = .resid, data = aschool) + geom_hline(yintercept = 0)
```
```{r }

# WOMEN'S LABOR FORCE
lmlabor <- lme(fixed = log_lifeexp ~ log_wlabor, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmlabor)

alabor <- augment(lmlabor)
qplot(x = .fitted, y = .resid, data = alabor) + geom_hline(yintercept = 0)
```
```{r }
# VOICE AND ACCOUNTABILITY
lmvoice <- lme(fixed = Life_Exp ~ Voice_Accountability, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmvoice)

avoice <- augment(lmvoice)
qplot(x = .fitted, y = .resid, data = avoice) + geom_hline(yintercept = 0)
```
```{r }
# INTERNET USE
lminternet <- lme(fixed = log_lifeexp ~ log_internet, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lminternet)

ainternet <- augment(lminternet)
qplot(x = .fitted, y = .resid, data = ainternet) + geom_hline(yintercept = 0)
```
```{r }
# POPULATION DENSITY
lmdensity <- lme(fixed = log_lifeexp ~ log_pop, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
summary(lmdensity)

adensity <- augment(lmdensity)
qplot(x = .fitted, y = .resid, data = adensity) + geom_hline(yintercept = 0)
```

```{r}
lmall <- lme(fixed = log_lifeexp ~ Fossil_Fuels + log_co2 + Overweight + log_fertility + GDP + log_pop + School_Enrollment + log_wlabor + log_internet + log_beds + Government_Effectiveness + log_wparliament + Political_Stability + Rule_Law + Voice_Accountability, 
              random = ~ 1 | Country,
              correlation = corAR1(form = ~ Year),
              data = esg_logs)
tidy(lmall)
plot(lmall)
```
# Examining Multicollinearity 

```{r}
vif(lmall)
```

# F-Tests
```{r}
broom.mixed::glance(lmall)
a_all <- augment(lmall)

sse_f <- sum(a_all$.resid^2)
df_f <- nrow(esg_logs) - 16 ## four parameters
sse_r <- sum((esg_logs$log_lifeexp - mean(esg_logs$log_lifeexp))^2)
df_r <- nrow(esg_logs) - 1 ## one parameter (beta_0)

F_star <- ((sse_r - sse_f) / (df_r - df_f)) / (sse_f / df_f)
F_star

pf(q = F_star, df1 = df_r - df_f, df2 = df_f, lower.tail = FALSE)
```

## Final Project Report

Since 2019 The Word Bank publishes a Sovereign ESG Data Framework Environment, Social, and Governance. This framework is divided into 3 pillars (Environmental, Social and Governance) which is based on their 17 sustainable development goals. The purpose is to help potential investors and shed light on countries’ sustainability performance. As for their design study, they collect data from different sources but their main data source is  the World Bank.We are interested in studying the associations that exist between the variables across these themes; specifically, we are interested in looking at expected lifespan as the response variable in our multiple regression model. We specifically are interested in seeing how the variables impact life expectancy across groups of people, such as gender, development status, and income, regions of the world including Latin America, European Union, Central Asia, and East Asia and the Pacific, and over a time span of 5 years. 

The variables we are interested in analyzing to study their association to life expectancy include: (1)  CO2 emissions (metric tons per capita) , (2) prevalence of overweight (% of adults), (3) hospital beds (per 1000 people) , (4) fertility rate (births per women), (5) government effectiveness (estimate), (6) portion of seats held by women in national parliaments (%), (7) political stability and absence of violence/terrorism (estimate), (8) GDP (annual growth %), (9) school enrollment primary and secondary and gender parity index (gross) , (10) ratio of female to male labor force participation rate (%), (11) voice and accountability (estimate), (12) population density (people per sq. km of land area)  and (13) fossil fuel (% of total), (14) Rule of Law, (15) Internet Usage. We are choosing these variables to explore different interests across group members and to explore possible associations among different categories of indicators (i.e. the impact of climate change on life expectancy). The sample size is 1,081. We wish to examine different variables that are not traditionally associated with life expectancy or social determinants of health. 

## Verification of Model Assumptions

The first assumption of the model of independence is not met because there are correlations of data points within each country over time. To remedy this, we used a linear mixed effect model that accounted for the correlation within countries. The assumptions of constant variance and linearity were assessed and met after performing residual analysis and then necessary log transformations on our response variable and predictor variables as needed. Furthermore, we assessed correlation between predictor variables by using the variance inflation factor that determined that the VIF of each variable was very close to 1 and there was no hidden multicollinearity between variables. 

## Statistical Methods

The first statistical method used was the overall F-test. The very low p-value of 8.303119e-130 provided evidence that at least one variable was very relevant to our data. The second method was transformations. This remedied curvature and nonconstant variance, but we ran into issues with many negative values that prevented the ability to do a log transformation. The third statistical method was using simple linear regression of individual variables, which gave us the ability to see how the variables interacted with our response variable and identified the necessary transformations. We also performed residual analysis and variance inflation factor as stated above. 

## Final Regression Models

The following is the model used for our Multiple Linear Regression model: 
log(Predicted life expectancy) = 3.89 - .00009* Fossil Fuels - .00007* log(CO2) + .005* Overweight - .0059* log(Fertility) - .000038* GDP + .024* log(Population Density) + .028* School Enrollment - .00075* log(Women’s Labor Participation Rate) + .0035* log(Internet Usage) + .0019* log(Hospital Beds) + .0013* Government Effectiveness + .000053* log(Women’s Participation in Parliament) + .00075* Political Stability + .0027* Rule of Law - .00038* Voice and Accountability.

Transformation was conducted by applying log to the response variable, life expectancy, and some of the predictor variables listed in the model. 

